### 2018-07-16 BATexFiles/Bibliography.tex
added some files that were not correctly tracked

\begin{thebibliography}{}
\bibitem{LifschitzASP}
Vladimir Lifschitz
\textit{Answer Set Planning}
ICLP, 1999.

\bibitem{Lifschitz2002AnswerSP}
Vladimir Lifschitz
\textit{Answer set programming and plan generation}
Artificial Intelligence, 2002, Vol. 138, p.39-54.

\bibitem{LifschitzWhatisASP}
Vladimir Lifschitz
\textit{What is Answer Set Programming?}
AAAI, 2008.

\bibitem{Gebser2013RicochetRA}
Martin Gebser, Holger Jost, Roland Kaminski, Philipp Obermeier, Orkunt Sabuncu, Torsten Schaub, Marius Thomas Lindauer
\textit{Ricochet Robots: A Transverse ASP Benchmark}
LPMNR, 2013.

\bibitem{BoteaHeuristicsVsPlanning}
Adi Botea, Martin Müller, Jonathan Schaeffer
\textit{Using Abstraction for Planning in Sokoban}
Computers and Games, 2002.

\bibitem{Frisch2006PropagationAF}
Alan M. Frisch, Brahim Hnich, Zeynep Kiziltan, Ian Miguel, Toby Walsh
\textit{Propagation algorithms for lexicographic ordering constraints}
Artificial Intelligence, 2006, Vol. 170, p.803-834.

\bibitem{Dor1999SOKOBANAO}
Dorit Dor, Uri Zwick
\textit{SOKOBAN and other motion planning problems}
Computational Geometry, 1999, Vol.13, p.215-228.

\bibitem{Froleyks2017UsingAA}
Nils Christian Froleyks and Tom{\'a}s Balyo
\textit{Using an Algorithm Portfolio to Solve Sokoban}
SOCS, 2017.
\end{thebibliography}